Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-04-21 draft.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/beta_A_Qerf1.1_corner_All.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$. Displayed point estimates represent the median of the posterior samples.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm}}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/beta_A_tp_Qerf1.1_corner.png}
    \caption{Posterior distributions for DTD parameters $\beta$, $A$, and $t_{\mathrm{p}}$. We display the log of $A$ for clarity. Displayed point estimates represent the median of the posterior samples.
    %The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm_tp}}
\end{figure}

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic \citep{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} = 1.3$ for the chains of those two parameters.

The joint and marginal posterior distributions for the fit parameters are found in Figs. \ref{fig:corner_beta_norm} and \ref{fig:corner_beta_norm_tp} for the fixed and free $t_{\mathrm{p}}$ fits respectively, and the posterior means and standard deviations are summarised in Table \ref{tab:dtd_results}.  Fig. \ref{fig:corner_beta_norm} shows that there is a correlation between $\beta$ and $A$ which is not seen in the fit with $t_{\mathrm{p}}$ as a free parameter. This is likely caused by $\beta$ instead correlating with $t_{\mathrm{p}}$, and being more poorly constrained in the three-parameter fit. In Fig. \ref{fig:rate_fitted} we show the SN rate per galaxy predicted by the model assuming the best fitting DTD parameters from Table \ref{tab:dtd_results} compared to the data as presented in Section \ref{sec:rates}. Visually, the model provides a good fit across a wide range of stellar mass. The model clearly diverges from the simple log-linear fit, describing better the enhanced rate around $10^{10}$ M$_{\odot}$ and suppression at high masses. This improvement provides evidence to support the DTD*SFH model. However, the model prediction still diverges from the data at high stellar masses ($>10^{11.25}$ M$_{\odot}$), suggesting that these data cannot be explained by our model, but are driven by further processes that we have not included. We discuss these in Section \ref{subsec:discussion}.

@@

The majority of works in the field have constrained the DTD power-law slope, and many also estimate its normalisation. Conversely, few have attempted to constrain the prompt time $t_{\mathrm{p}}$. In most cases, $t_{\mathrm{p}}$ has been fixed at some fiducial value such as 40~Myr, which is derived from the lifetime of 8~M$_{\odot}$ stars. \citet{Castrillo2020} included $t_{\mathrm{p}}$ (which they denote $\Delta$) in their fit, and find a value of $50_{-35}^{+100}$~Myr which is consistent with both 40~Myr as well as our best fitting value of $67_{-23}^{+43}$~Myr for the three-parameter fit and $47_{-7}^{+8}$~Myr with $\beta$ and $A$ fixed. \citet{Heringer2019} also present fits with varying prompt times, although they don't fit for the parameter itself. They find that varying $t_{\mathrm{p}}$ changes the recovered normalisation, which is consistent with the degeneracy seen in our results (Fig. \ref{fig:corner_beta_norm_tp}). Overall, we find results for $t_{\mathrm{p}}$ that are consistent with the fiducial value of 40~Myr and adopt this value for the remainder of the analysis.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_all_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$. Displayed point estimates represent the median of the posterior samples.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm}}
\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[width=.5\textwidth]{figs/fit_dtd_A_beta_tp_corner.png}
%    \caption{Posterior distributions for DTD parameters $\beta$, $A$, and $t_{\mathrm{p}}$. We display the log of $A$ for clarity. Displayed point estimates represent the median of the posterior samples.
%    %The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
%    \label{fig:corner_beta_norm_tp}}
%\end{figure}

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} = 1.3$ for the chains of those two parameters. In subsequent analyses we adopt as our fiducial model that with fixed $t_{\mathrm{p}} =0.04$~Gyr.

The joint and marginal posterior distributions for the fit parameters are found in Fig. \ref{fig:corner_beta_norm}, and the posterior means and standard deviations are summarised in Table \ref{tab:dtd_results}.  Fig. \ref{fig:corner_beta_norm} shows that there is a mild correlation between $\beta$ and $A$. In Fig. \ref{fig:rate_fitted} we show the SN rate per galaxy predicted by the model assuming the best fitting DTD parameters from Table \ref{tab:dtd_results} compared to the data as presented in Section \ref{sec:rates}. Visually, the model provides a good fit across a wide range of stellar mass. The model clearly diverges from the simple log-linear fit, describing better the enhanced rate around $10^{10}$ M$_{\odot}$ and suppression at high masses. This improvement provides evidence to support the DTD*SFH model. However, the model prediction still diverges from the data at high stellar masses ($>10^{11.25}$ M$_{\odot}$), suggesting that these data cannot be explained by our model, but are driven by further processes that we have not included. We discuss these in Section \ref{subsec:discussion}.

@@

The majority of works in the field have constrained the DTD power-law slope, and many also estimate its normalisation. Conversely, few have attempted to constrain the prompt time $t_{\mathrm{p}}$. In most cases, $t_{\mathrm{p}}$ has been fixed at some fiducial value such as 40~Myr, which is derived from the lifetime of 8~M$_{\odot}$ stars. \citet{Castrillo2020} included $t_{\mathrm{p}}$ (which they denote $\Delta$) in their fit, and find a value of $50_{-35}^{+100}$~Myr which is consistent with both 40~Myr. While we are unable to fit $t_{\mathrm{p}}$ simultaneously with $\beta$ and $A$, by fixing $\beta=-1$ and $\log(A)=-12.75$ we find $t_{\mathrm{p}} =47_{-7}^{+8}$~Myr, which is consistent with the fiducial value of $40$~Myr. \citet{Heringer2019} also present fits with varying prompt times, although they don't fit for the parameter itself. They find that varying $t_{\mathrm{p}}$ changes the recovered normalisation, which is consistent with the degeneracy encountered in our fits. Overall, we conclude that there is no strong evidence for a value of $t_{\mathrm{p}}$ that is different to the fiducial value of 40~Myr and adopt this value for the remainder of the analysis.
